Sketch of DP - Bayesian Inference

### 2018-07-17 main-sketch.tex
typos

\begin{abstract}
Bayesian inference is a statistical method which allows to derive a
\emph{posterior} distribution over a parameter, starting from a
\emph{prior} distribution and observed data.  In a setting where the
data must be protected, different approaches have been taken to make
this process differentially private. \citet{dimitrakakis2014robust},
and \citet{wang2015privacy}, for instance, proved that, under specific
conditions, sampling from the posterior distribution is already
differentially private. Other works,
e.g. \cite{zhang2016differential}, \cite{foulds2016theory}, designed
differentially private mechanisms that output a representation of the
full posterior distribution. Also, accuracy of the mechanisms was
measured using metrics on the space of the numeric parameters of the
posterior distribution, e.g. $\ell_1$-norm .

In this work we present a new differentially private algorithm to
compute the posterior in a Beta-Binomial system and in Dirichlet-Multinomial system.
The noise is calibrated w.r.t. a smooth
sensitivity computed with a metric on probability distributions; we
measured the accuracy of the process by using a metric over
probability measures. We compared the accuracy of this approach with
the one based on $\ell_1$-norm. Experimental results show that when
the data size is small or when the parameters of the prior
distribution are large, the former outperforms the latter.

@@

\section{Bayesian Inference Background}
\label{sec_background}
Given a prior belief $\Pr(\theta)$ on some parameter $\theta$,
and given an observation $\dataobs$,
the posterior distribution on $\theta$ given $\dataobs$ is computed as:
\[
  \Pr(\theta | \dataobs) = \frac{\Pr(\dataobs | \theta) \cdot \Pr(\theta)}{\Pr(\dataobs)}
\]
where the expression  $\Pr(\dataobs | \theta)$ denotes the \emph{likelihood} of $\theta$ when
$\dataobs$ is observed. Since we consider $\dataobs$ to be fixed the likelihood is a function of $\theta$.
For the same reason $\Pr(\dataobs)$ is a constant independent of $\theta$.
Usually in statistics the prior distribution $\Pr(\theta)$ is chosen so that it represents
the initial belief on $\theta$. That is, when no data has been observed. In practice though,
prior distributions and likelihood functions are usually chosen so that the posterior
belongs to the same \emph{family} of distributions. In this case we say that the prior
is conjugate of the likelihood function. Using conjugate priors simplify calculcations and allows
for inference to be performed in a recursive fashion over the data.
In this work we will consider a specific instance of Bayesian inference and one of its generalizations.
Specifically, we will consider the situation where $\theta$ represents the parameter --informally called \emph{bias}--
of a Bernoulli distributed random variable, and its immediate generalization where the parameter $\vtheta$ represents
the the vector of parameters of a categorical distributed random variable.
In the former case the prior distribution over $\theta\in [0,1]$ is going to be a beta distribution, $\betad(\alpha, \beta)$, with parameters
$\alpha,\beta\in\mathbb{R}^{+}$, and with p.d.f:
\[
  \Pr(\theta)\equiv \frac{\theta^{\alpha} (1- \theta)^{\beta}}{\betaf(\alpha,\beta)}
\]
where $\betaf(\cdot,\cdot)$ is the beta function.
The data $\dataobs$ will be a sequence of $n\in\mathbb{N}$ 0/1 values, that is $\dataobs=\ll x_1,\dots x_n \gg, x_i\in\{0,1\}$, and the likelihood function is:
\[
  \Pr(\dataobs | \theta)\equiv \theta^{\Delta \alpha}(1-\theta)^{n - \Delta \alpha}
\]
where $\Delta \alpha = \displaystyle\sum_{i=1}^{n}x_i$.
From this it can easily be derived that the posterior distribution is:
\[
  \Pr(\theta|\dataobs)=\betad(\alpha + \Delta \alpha,\beta + n - \Delta \alpha)
\]
In the latter case the prior distribution over $\vtheta\in [0,1]^{k}$ is given by a Dirichelet distribution, $\dirichlet(\valpha)$, for $k\in\mathbb{N}$,
and $\valpha\in(\mathbf{R^{+}})^{k}$, with p.d.f:
\[
  \Pr(\vtheta)\equiv\frac{1}{\mbetaf(\valpha)}\cdot \displaystyle\prod_{i=1}^{k}{\theta_i^{\alpha_i-1}}
\]
where $\mbetaf(\cdot)$ is the generealised beta function.
The data $\dataobs$ will be a sequence of $n\in\mathbb{N}$ values coming from a universe $\datauni$, such that $\mid\datauni \mid=k$.
The likelihood function will be:
\[
  \Pr(\dataobs|\vtheta)\equiv \displaystyle\prod_{a_i\in\datauni}\theta_{i}^{\Delta \alpha_i},
\]
where $\Delta \alpha_i=\displaystyle\sum_{i=1}^{n}\iverson{x_i=a_i}$, where in $\iverson{\cdot}$ we use
Iverson bracket notation. The posterior distribution over $\vtheta$ turns out to be
\[
  \Pr(\vtheta|\dataobs)=\dirichlet(\valpha+\Delta \valpha). 
\]
where $+$ denotes the component wise  sum of vectors of reals.

@@

\noindent Also, in previous works, the accuracy has been measured again with respect to
$\ell_1$ norm. That is, a lower bound has been given to
\[
  \Pr[|\alpha - \tilde\alpha| + |\beta - \tilde\beta |\geq \gamma ]
\]
where $(\alpha, \beta), (\tilde\alpha,\tilde\beta)$ are as  defined above.
In this work we will use a metric based on a different norm to compute the sensitivity
and provide guarantees on the accuracy. In particular we will consider a metric
over probability measures and not over the parameters that represent them.
Specifically, we will use the Hellinger distance $\hellinger(\cdot,\cdot)$, which for two beta distributions
$\boldsymbol{\beta}_1=\betad(\alpha_1, \beta_1),$ and $\boldsymbol{\beta}_2=\betad(\alpha_2, \beta_2)$ is equal to
\[
  \hellinger(\boldsymbol{\beta}_1, \boldsymbol{\beta}_2)\equiv
  \sqrt{1 - \frac{\betaf(\frac{\alpha_1 + \alpha_2}{2}, \frac{\beta_1 + \beta_2}{2})}{\sqrt{\betaf(\alpha_1,\beta_1)\betaf(\alpha_2,\beta_2)}}}
\]
The same change of metric will be applied to the experimental accuracy guarantees.
% \section{Technical Problem Statement}
% \label{sec_bayesInfer}
% The posterior belief about parameters $\xis$ is inferred from the
% group of sensitive data. Because it is usually published in form of a
% distribution, for example, $\betad(\alpha',\beta'), \cdots,$ itself
% can relieve a lot of information about sensitive data. So, the
% sensitive information we want to protect here is the posterior belief,
% i.e., $Pr(\xis | \dataobs)$, rather than just a sample drew from
% it\cite{foulds2016theory}. Then it will be better if we can use a
% distance over two distributions to measure the accuracy than a
% distance over just parameters or values when we are protecting a
% distribution, for instance, Hellinger distance in our paper. When we
% measure the accuracy between true posterior and protected posterior
% using Hellinger distance, it would be better we can have a protection
% mechanism based on Hellinger distance. However, current mechanisms are
% mainly based on $l_1$ norm by Laplace mechanism
% \cite{zhang2016differential} \cite{xiao2012bayesian}. That's why we
% design our exponential mechanism over Hellinger distance to protect
% the whole posterior distribution.

@@

We can now define the mechanism $\hexpmech$ which,
given in input a sequence of observations $\dataobs$, and $\epsilon>0, \delta>0$,
produces an element $r$ in $\betaset$ with probability:

\begin{equation*}
\underset{z \thicksim \hexpmech}{\Pr}[z=r] = \frac {exp(\frac{-\epsilon\cdot\hellinger(\bysinfer(\dataobs),r)}{2\cdot S_\beta(\dataobs)})}
{\displaystyle\sum_{r\in\betaset} exp(\frac{-\epsilon\cdot\hellinger(\bysinfer(\dataobs),r)}{2\cdot S_{\beta}(\dataobs)})}
\end{equation*}
this mechanism is based on the basic exponential mechanism \cite{talwar}, with $\betaset$ as range and $\hellinger(\cdot,\cdot)$ as scoring function.
The difference is that in this mechanism we don't calibrate the noise w.r.t to the global sensitivity of the scoring function but w.r.t to the smooth sensitivity
-- \citet{nissim2007smooth}-- of $\hellinger(\bysinfer(\dataobs), \cdot)$, which is :
\[
   S_{\beta}(\dataobs)=\max_{\dataobs' \neq \dataobs, \dataobs' \in \{0,1\}^{n}}\bigg \{\Delta_{l}\bigg (\hellinger(\bysinfer(\dataobs'),\cdot)\bigg )\cdot e^{-\gamma\cdot d(\dataobs, \dataobs')}\bigg\}
 \]

>>>>

\begin{abstract}
Bayesian inference is a statistical method which allows to derive a
\emph{posterior} distribution over a parameter, starting from a
\emph{prior} distribution and observed data.  In a setting where the
data must be protected, different approaches have been taken to make
this process differentially private. \citet{dimitrakakis2014robust},
and \citet{wang2015privacy}, for instance, proved that, under specific
conditions, sampling from the posterior distribution is already
differentially private. Other works,
e.g. \cite{zhang2016differential}, \cite{foulds2016theory}, designed
differentially private mechanisms that output a representation of the
full posterior distribution. Also, accuracy of the mechanisms was
measured using metrics on the space of the numeric parameters of the
posterior distribution, e.g. $\ell_1$-norm.

In this work we present a new differentially private algorithm to
compute the posterior in a Beta-Binomial system and in Dirichlet-Multinomial system.
The noise is calibrated w.r.t. a smooth
sensitivity computed with a metric on probability distributions; we
measured the accuracy of the process by using a metric over
probability measures. We compared the accuracy of this approach with
the ones based on $\ell_1$-norm. Experimental results show that when
the data size is small or when the parameters of the prior
distribution are large, the former outperforms the latter.

@@

\section{Bayesian Inference Background}
\label{sec_background}
Given a prior belief $\Pr(\theta)$ on some parameter $\theta$,
and given an observation $\dataobs$,
the posterior distribution on $\theta$ given $\dataobs$ is computed as:
\[
  \Pr(\theta | \dataobs) = \frac{\Pr(\dataobs | \theta) \cdot \Pr(\theta)}{\Pr(\dataobs)}
\]
where the expression  $\Pr(\dataobs | \theta)$ denotes the \emph{likelihood} of $\theta$ when
$\dataobs$ is observed. Since we consider $\dataobs$ to be fixed the likelihood is a function of $\theta$.
For the same reason $\Pr(\dataobs)$ is a constant independent of $\theta$.
Usually in statistics the prior distribution $\Pr(\theta)$ is chosen so that it represents
the initial belief on $\theta$. That is, when no data has been observed. In practice though,
prior distributions and likelihood functions are usually chosen so that the posterior
belongs to the same \emph{family} of distributions. In this case we say that the prior
is conjugate of the likelihood function. Using conjugate priors simplify calculcations and allows
for inference to be performed in a recursive fashion over the data.
In this work we will consider a specific instance of Bayesian inference and one of its generalizations.
Specifically, we will consider the situation where $\theta$ represents the parameter --informally called \emph{bias}--
of a Bernoulli distributed random variable, and its immediate generalization where the parameter $\vtheta$ represents
the the vector of parameters of a categorical distributed random variable.
In the former case the prior distribution over $\theta\in [0,1]$ is going to be a beta distribution, $\betad(\alpha, \beta)$, with parameters
$\alpha,\beta\in\mathbb{R}^{+}$, and with p.d.f:
\[
  \Pr(\theta)\equiv \frac{\theta^{\alpha} (1- \theta)^{\beta}}{\betaf(\alpha,\beta)}
\]
where $\betaf(\cdot,\cdot)$ is the beta function.
The data $\dataobs$ will be a sequence of $n\in\mathbb{N}$ binary values, that is $\dataobs=\langle x_1,\dots x_n \rangle, x_i\in\{0,1\}$, and the likelihood function is:
\[
  \Pr(\dataobs | \theta)\equiv \theta^{\Delta \alpha}(1-\theta)^{n - \Delta \alpha}
\]
where $\Delta \alpha = \displaystyle\sum_{i=1}^{n}x_i$.
From this it can easily be derived that the posterior distribution is:
\[
  \Pr(\theta|\dataobs)=\betad(\alpha + \Delta \alpha,\beta + n - \Delta \alpha)
\]
In the latter case the prior distribution over $\vtheta\in [0,1]^{k}$ is given by a Dirichelet distribution, $\dirichlet(\valpha)$, for $k\in\mathbb{N}$,
and $\valpha\in(\mathbf{R^{+}})^{k}$, with p.d.f:
\[
  \Pr(\vtheta)\equiv\frac{1}{\mbetaf(\valpha)}\cdot \displaystyle\prod_{i=1}^{k}{\theta_i^{\alpha_i-1}}
\]
where $\mbetaf(\cdot)$ is the generealised beta function.
The data $\dataobs$ will be a sequence of $n\in\mathbb{N}$ values coming from a universe $\datauni$, such that $\mid\datauni \mid=k$.
The likelihood function will be:
\[
  \Pr(\dataobs|\vtheta)\equiv \displaystyle\prod_{a_i\in\datauni}\theta_{i}^{\Delta \alpha_i},
\]
with $\Delta \alpha_i=\displaystyle\sum_{j=1}^{n}\iverson{x_j=a_i}$, where in $\iverson{\cdot}$ we use
Iverson bracket notation. Denoting by $\Delta\valpha$ the vector $\langle  \Delta\alpha_1,\dots \Delta\alpha_k \rangle$ the posterior distribution over $\vtheta$ turns out to be
\[
  \Pr(\vtheta|\dataobs)=\dirichlet(\valpha+\Delta \valpha). 
\]
where $+$ denotes the component wise  sum of vectors of reals.

@@

\noindent Also, in previous works, the accuracy has been measured again with respect to
$\ell_1$ norm. That is, an upper bound has been given to
\[
  \Pr[|\alpha - \tilde\alpha| + |\beta - \tilde\beta |\geq \gamma ]
\]
where $(\alpha, \beta), (\tilde\alpha,\tilde\beta)$ are as  defined above.
In this work we will use a metric based on a different norm to compute the sensitivity
and provide guarantees on the accuracy. In particular we will consider a metric
over probability measures and not over the parameters that represent them.
Specifically, we will use the Hellinger distance $\hellinger(\cdot,\cdot)$. Given two beta distributions
$\boldsymbol{\beta}_1=\betad(\alpha_1, \beta_1),$ and $\boldsymbol{\beta}_2=\betad(\alpha_2, \beta_2)$ the following equality holds 
\[
  \hellinger(\boldsymbol{\beta}_1, \boldsymbol{\beta}_2)=
  \sqrt{1 - \frac{\betaf(\frac{\alpha_1 + \alpha_2}{2}, \frac{\beta_1 + \beta_2}{2})}{\sqrt{\betaf(\alpha_1,\beta_1)\betaf(\alpha_2,\beta_2)}}}
\]
The same change of metric will be applied to the experimental accuracy guarantees.
% \section{Technical Problem Statement}
% \label{sec_bayesInfer}
% The posterior belief about parameters $\xis$ is inferred from the
% group of sensitive data. Because it is usually published in form of a
% distribution, for example, $\betad(\alpha',\beta'), \cdots,$ itself
% can relieve a lot of information about sensitive data. So, the
% sensitive information we want to protect here is the posterior belief,
% i.e., $Pr(\xis | \dataobs)$, rather than just a sample drew from
% it\cite{foulds2016theory}. Then it will be better if we can use a
% distance over two distributions to measure the accuracy than a
% distance over just parameters or values when we are protecting a
% distribution, for instance, Hellinger distance in our paper. When we
% measure the accuracy between true posterior and protected posterior
% using Hellinger distance, it would be better we can have a protection
% mechanism based on Hellinger distance. However, current mechanisms are
% mainly based on $l_1$ norm by Laplace mechanism
% \cite{zhang2016differential} \cite{xiao2012bayesian}. That's why we
% design our exponential mechanism over Hellinger distance to protect
% the whole posterior distribution.

@@

We can now define the mechanism $\hexpmech$ which,
given in input a sequence of observations $\dataobs$, $\epsilon>0$ and $\delta>0$,
produces an element $r$ in $\betaset$ with probability:

\begin{equation*}
\underset{z \thicksim \hexpmech}{\Pr}[z=r] = \frac {exp(\frac{-\epsilon\cdot\hellinger(\bysinfer(\dataobs),r)}{2\cdot S_\beta(\dataobs)})}
{\displaystyle\sum_{r\in\betaset} exp(\frac{-\epsilon\cdot\hellinger(\bysinfer(\dataobs),r)}{2\cdot S_{\beta}(\dataobs)})}
\end{equation*}
this mechanism is based on the basic exponential mechanism \cite{talwar}, with $\betaset$ as range and $\hellinger(\cdot,\cdot)$ as scoring function.
The difference is that in this mechanism we don't calibrate the noise w.r.t to the global sensitivity of the scoring function but w.r.t to the smooth sensitivity $S_{\beta}(\dataobs)$
-- \citet{nissim2007smooth}-- of $\hellinger(\bysinfer(\dataobs), \cdot)$. The smooth sensitivity is computed as follows:
\[
   S_{\beta}(\dataobs)=\max_{\dataobs' \neq \dataobs, \dataobs' \in \{0,1\}^{n}}\bigg \{\Delta_{l}\bigg (\hellinger(\bysinfer(\dataobs'),\cdot)\bigg )\cdot e^{-\gamma\cdot d(\dataobs, \dataobs')}\bigg\}
 \]
